Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-24 paper/high-assurance-contracts.tex
forgot unwritten file
wellformedness discussion (partial)

\subsection{Code contracts}
\label{code-contracts}

Generally, AGREE specifications do not describe the computation that a
component performs. This is entirely by design: AGREE is intended to
reason about component behavior solely at the specification
level. However, the syntax of AGREE specifications provides enough
expressiveness to support the notion of a \emph{code contract}: a
contract from which an implementation can be extracted. First we must
discuss a class of guarantees---\emph{output guarantees}---which
determine the values on all output ports of a component.

@@

The semantics of Section \ref{agree-semantics} supports a formal
connection between the original contract (and verification results of
the AGREE model checker being run on it), and code generated from
specifications of leaf level cyber-components. It also provides the
root meaning at the base of a chain of translation steps moving from
AGREE specifications to CakeML executables.

The first step in the chain maps the contract to a code-focused
representation. Assume given code contract $(I,O,A,\mathit{Eqs} \cup G \cup
P,\emptyset)$, environment $E$, and time $t$. The evaluation of $\mathit{Eqs}$
followed by the evaluation of $G$ results in a new environment $E'$
where the stream values for state variables and outputs have been
computed for time $t$. The step from $E$ to $E'$ is one full cycle in
the repeated evaluation of the component.

\begin{definition}[Evaluation]
We overload existing notation and write the evaluation of
$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
$\sem{\mathit{Eqs}\cdot G}^E_t$. Each $v = e \in {\Eqs}$ is treated as
an environment transformation
\[
 \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
\]
modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
{\Eqs} is evaluated by folding the transformation left-to-right
through it. The transformation is also applied to the list of output
guarantees $G$, computing the values on output ports. The details are
omitted, being a bit messy because of the different kinds of output
ports available.
\end{definition}

@@

How does this notion of correctness integrate with the verification
conditions addressed by model checking? In \emph{contract
correctness}, the model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intution, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the tepid
bathwater of imperative program verification. In order to use the code
contract result in system contract verifcation conditions, one has to
show that the code contract result implies the system verification
condtion.

\subsubsection*{Wellformedness}
The evaluation semantics given above is based on the stream
computation model, where values `deep in the past' can be accessed in
computations. However, the conventional program languages we target
target do not support such a view. Thus code generation is only
defined for a class of \emph{wellformed} equations.

\begin{definition}[Wellformedness]
\end{definition}

>>>>

Generally, AGREE contracts do not describe the computation that a
component performs. This is entirely by design: AGREE is intended to
reason about component behavior solely at the specification
level. However, the syntax of AGREE provides enough expressiveness to
support the notion of a \emph{code contract}: a contract from which an
implementation can be extracted. First we must discuss a class of
guarantees---\emph{output guarantees}---which determine the values on
all output ports of a component.

@@

The semantics of Section \ref{agree-semantics} supports a formal
connection between the original contract---and verification results of
the AGREE model checker being run on it---and code generated from
specifications of leaf level cyber-components. It also provides the
root meaning at the base of a chain of translation steps moving from
an AGREE contract to a CakeML executables.

The first step in the chain maps the contract to a code-focused
representation. Assume given a code contract
$(I,O,A,\mathit{Eqs} \cup G \cup P,\emptyset)$, environment $E$, and
time $t$. The \emph{evaluation} of $\mathit{Eqs}$ followed by the
evaluation of $G$ results in a new environment $E'$ where the stream
values for state variables and outputs have been computed for time
$t$. The step from $E$ to $E'$ is one full cycle in the repeated
evaluation of the component.

\begin{definition}[Evaluation]
We overload existing notation and write the evaluation of
$\mathit{Eqs}$ and then $G$ in environment $E$ at time $t$ as
$\sem{\mathit{Eqs}\cdot G}^E_t$. The evaluation of $v = e \in {\Eqs}$
is an environment transformation
\[
 \sem{v = e}^E_t = E[v(t) \mapsto\sem{e}^E_t] \ .
\]
modifying $E$ so that stream $v$ has value $\sem{e}^E_t$ at time $t$.
The list {\Eqs} is evaluated by folding the transformation left-to-right
through it. The transformation is similarly applied to the list of output
guarantees $G$, computing the values on output ports. The details are
omitted, being a bit messy because of the different kinds of output
ports available.
\end{definition}

@@

How does this notion of correctness integrate with the verification
conditions addressed by model checking? In \emph{contract
correctness}, the model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intuition, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the tepid
bathwater of imperative program verification. For example, at this
point one could apply a verification condition generator to help break
down the proof obligation. In order to use the code contract result in
system contract verifcation conditions, one has to show that the code
contract result implies the system verification condtion.

\subsubsection*{Wellformedness}
The evaluation semantics given above is based on the stream
computation model, where values arbitrarily `deep in the past' can be
accessed in computations by use of $\konst{pre}(-)$. However, the
conventional program languages we target do not support such a
view. We are currently examining transformations that compile such
deep temporal accesses away by introducing intermediate
variables. Avoiding discussion of the algorithm, we focus on the
notion of \emph{wellformed} equations that needs to hold. The guiding
intuition is that an \emph{implementable} {\Eqs} can allow access at
most one step in the past, and accesses of a past value for a variable
are only possible until the variable is updated. In lieu of a formal
definition, we give a few examples.

\begin{definition}[Wellformedness]

\end{definition}
